---
title: "PS270 Final Project"
author: "Evelyn Berg"
description: "My final project"
output:
  distill::distill_article:
    self_contained: false
---


## Milestone 2

```{r}
library(haven)
library(ggplot2)
library(tidyverse)
library(dplyr)
```


```{r}
data <- read_dta("Data.dta")
head(data)
```


My research question is how does  electoral integrity after national-level elections vary across the world? In this project, I will look at values representing electoral integrity for a number of different countries. I hypothesize that electoral integrity will tend to be higher in countries in the more developed countries. Moreover, I expect that it will be higher in the most democratic countries in the data set. I expect these findings because stability is very conducive to government legitimacy. My data is taken from 170 countries across 643 elections. My explanatory variables of interest are the different countries studied in the data. My outcome variable of interest is electoral integrity. This is measured through an index of 64 indicators that make up 11 categories. In my data, an observed pattern of higher scores in more developed countries would support my hypothesis. On the other hand, higher scores in less developed countries would disprove my hypothesis. 


## Milestone 3 (Revised Visualization)

```{r}
data |>
  filter(region9 == 3, !is.na(PEI_add_original)) |>

ggplot(mapping = aes(x = development,
                     y = PEI_add_original)) +
  geom_point(mapping = aes(color = country), size = 2, alpha = 0.5) +
  geom_smooth(method = "lm", se = FALSE,  formula = y ~ x) +
  labs(x = "World Bank Development Index",
       y = "Original Perceptions of Electoral Integrity (PEI)"
       , title = "The Effect of Development on PEI") 

```

## Milestone 4 (Analysis & Results Via 2nd Visualization)

```{r}
data |>
  filter(region9 == 3) |>
  group_by(development) |>
  summarize(avg_PEI = mean(PEI_add_original, na.rm = TRUE)) |>
  ggplot(aes(x = factor(development), y = avg_PEI)) +
  geom_col() +
  labs(
    x = "Development Level",
    y = "Average PEI",
    title = "Average Electoral Integrity by Development Level in East Asia & the Pacific")
  data
```
According to my visualizations, there appears to be a positive relationship between development and electoral integrity. The line in the first graph has a positive slope, and both the scatter plot and the bar plot show that the highest levels of electoral integrity tend to come from the countries with the highest development scores. To answer my reseearch question with this evidence, it appears that perceptions of electoral integrity are better in more developed countries. 
